Constructing hierarchical time series through clustering: \\Is there an optimal way for forecasting?

### 2024-04-15 manuscript/titlepage.tex
submit to IJF

% \usepackage{natbib}
\usepackage{amsfonts,amsmath,bbm,bm,xcolor,booktabs,hyperref,amsthm}
\usepackage{geometry}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{subfig}
\geometry{a4paper,scale=0.8}
\usepackage{setspace}
\setstretch{1.5}
\let\code=\texttt
\let\proglang=\textsf
\setcounter{MaxMatrixCols}{20}
\usepackage{enumitem}

% Define a new counter for your research questions
\newcounter{researchquestion}
\setcounter{researchquestion}{0}

% Define the environment for the research questions
\newenvironment{researchquestion}{%
    \refstepcounter{researchquestion}%
    \par\medskip\noindent%
    \textbf{RQ\theresearchquestion.}~%
}{\medskip}

@@

% ADDING LINENUMBERS FOR REVIEWING:
% \usepackage{lineno}

@@

  \begin{abstract}

    Forecast reconciliation has attracted significant research interest in recent years, with most studies taking the hierarchy of time series as given. We extend existing work that uses time series clustering to construct hierarchies, with the goal of improving forecast accuracy, in three ways. First, we investigate multiple approaches to clustering, including not only different clustering algorithms, but also the way time series are represented and how distance between time series is defined. We find that cluster-based hierarchies lead to improvements in forecast accuracy relative to two-level hierarchies. Second, we devise an approach based on random permutation of hierarchies, keeping the structure of the hierarchy fixed, while time series are randomly allocated to clusters. In doing so, we find that improvements in forecast accuracy that accrue from using clustering do not arise from grouping together similar series but from the structure of the hierarchy. Third, we propose an approach based on averaging forecasts across hierarchies constructed using different clustering methods, that is shown to outperform any single clustering method. All analysis is carried out on two benchmark datasets and a simulated dataset. Our findings provide new insights into the role of hierarchy construction in forecast reconciliation and offer valuable guidance on forecasting practice. \\

    %the dynamics between ``grouping'' and ``structure'', which lead to an improved understanding of forecast reconciliation.\\

  \end{abstract}

  \begin{keyword}
  Forecast reconciliation \sep
  Hierarchical time series \sep
  Clustering \sep
  Hierarchy construction \sep
  Forecast combination
  \end{keyword}

@@

%\clearpage
\newpage
% \linenumbers

\section{Introduction}

Applications where some variables are aggregates of one another, or so-called \textit{hierarchical time series (HTS)}, are found in many forecasting problems ranging from supply chain management (\citealp{syntetosSupplyChainForecasting2016}) to tourism planning (\citealp{kourentzesCrosstemporalCoherentForecasts2019}), electrical load forecasting (\citealp{jeonProbabilisticForecastReconciliation2019}), and retail demand forecasting (\citealp{makridakisM5AccuracyCompetition2022}). In recent decades, there has been an increasing interest in hierarchical forecasting, primarily driven by the success of the optimal reconciliation framework (\citealp{hyndmanOptimalCombinationForecasts2011,wickramasuriyaOptimalForecastReconciliation2019, panagiotelisProbabilisticForecastReconciliation2023}). The original motivation for forecast reconciliation was to ensure forecasts are \textit{coherent}, that is they respect the aggregation constraints implied by the hierarchical structure. Coherent forecasts facilitate aligned decisions by agents acting upon different variables within the hierarchy. For example, consider a retail setting, where a warehouse manager supplies stock to individual store managers within their region. Forecasts could be incoherent when the warehouse manager forecasts low total demand while store managers forecast high demand, leading to supply shortages. Numerous case studies in the literature demonstrate that reconciliation approaches not only yield coherent forecasts but also enhance overall forecast performance (\citealp{AthanasopoulosForecastReconciliationReview2023}).

%The primary objective of hierarchical forecasting is to produce coherent forecasts for the series within a given HTS, which we refer to as \textit{original hierarchy}, ensuring that decision-makers at different levels of the hierarchy can analyze and make informed decisions based on aligned expectations for the future.
A limitation in the overwhelming majority of forecast reconciliation studies is that the structure of the hierarchy is taken as \textit{given}. This structure usually includes \textit{bottom level series}, an overall aggregate or \textit{top level series}, with various aggregation schemes used to construct \textit{middle level series}. Typically, middle levels are formed according to inherent attributes of the bottom-level series, such as geographical location, gender, product category, travel purpose, and others. We refer to this type of structure as the \textit{natural hierarchy}. While in some forecasting applications, decisions must be made with respect to the natural hierarchy, in other settings there might be some flexibility in determining how bottom levels are aggregated into middle levels. Moreover, a predefined natural hierarchy may not always exist in hierarchical settings, underscoring the importance of a hierarchical construction framework for forecasting purposes. It should be noted that very little attention has been paid to whether middle level series can be constructed in a way that leads to further improvements in forecast accuracy relative to a given natural hierarchy. To the best of our knowledge, only \cite{pangHierarchicalElectricityTime2018}, \cite{liForecastReconciliationApproach2019}, \cite{pangHierarchicalElectricityTime2022}, and \cite{matteraImprovingOutofSampleForecasts2023} have attempted to \textit{construct} middle level series in a data-driven way which ultimately improved forecast accuracy. However, all of these works use time series clustering to construct hierarchies in a manner that is somewhat ad hoc. Our work conducts a more thorough investigation of issues faced when  constructing hierarchical structures in forecast reconciliation. In particular, we address the following three research questions:%via time series clustering. %We also assess and disentangle the impact of

%This limitation is evident in the absence of any discussion regarding cluster-based approaches in the recent comprehensive review of forecast reconciliation by \cite{AthanasopoulosForecastReconciliationReview2023}. However, the following questions arise when pursuing superior forecast performance, motivating the investigation presented in this paper.

\begin{researchquestion} In terms of forecast performance, can the use of middle level series lead to improvement compared to a two-level hierarchy (consisting of only top and bottom time series)? If so, is it possible to construct hierarchies in a data-driven way that leads to further improvements in forecast accuracy?

%In terms of forecast performance, is the natural hierarchy better than a two-level hierarchy (consisting of only top and bottom time series)? 
%In terms of forecast performance, does the natural hierarchy outperform a two-level hierarchy of only top and bottom time series?
\end{researchquestion}
%\begin{researchquestion}
%    If the use of middle level series in the ``natural'' hierarchy can lead to improvement in forecast accuracy, is it possible to construct hierarchies in a data-driven way that leads to further improvements in forecast accuracy?
%\end{researchquestion}

%Two observations have been consistently demonstrated in empirical studies on forecast reconciliation. First, the accuracy of base forecasts, which are influenced by the individual time series characteristics within the hierarchy, critically affects the overall performance of reconciliation. For instance, \cite{panagiotelisForecastReconciliationGeometric2021} show that greater performance improvement can be achieved when bias correction for base forecasts is involved. 
%Under the reconciliation framework, improvements in accuracy for some series often come at the expense of others (see \citealp{pritulargaStochasticCoherencyForecast2021}).
To investigate these questions, we consider two widely used empirical HTS datasets; the first is Australian tourism demand, the second, cause-of-death mortality data.  Throughout the paper, all evaluations are carried out using the series common to all hierarchies; namely the top and bottom level series. In both datasets, we find that the natural hierarchy outperforms the two-level hierarchy, and data-driven hierarchy via clustering can further improve forecast performance compared to natural hierarchy.

The rationale behind the data-driven approach lies in grouping time series with similar patterns together, thereby creating middle-level series with enhanced signals and consequently, improved forecastability. Such arguments have been put forward by  \cite{pangHierarchicalElectricityTime2018}, \cite{liForecastReconciliationApproach2019}, \cite{pangHierarchicalElectricityTime2022}, and \cite{matteraImprovingOutofSampleForecasts2023}. They all demonstrate superior forecast performance from hierarchies constructed via clustering, relative to the natural hierarchy or the two-level hierarchy. However, these studies for the most part focus on a small number of (in some cases, a single) clustering techniques. In this paper, we take a more systematic approach by clustering time series using different representations (the original time series, forecast errors, features of both), different distance metrics (Euclidean, dynamic time warping), and different clustering paradigms ($k$-medioids, hierarchical). The models used to obtain base forecasts and the reconciliation method are fixed throughout the experiments. Using both empirical datasets, as well as a simulation study, we find evidence that constructing hierarchies via clustering can lead to improved forecasting performance, although the optimal clustering method depends on the dataset characteristics.% \textbf{We use the same base forecasting model and reconciliation method through this paper}.

While the idea behind time series clustering is intuitively appealing, the increased accuracy when using clustering-based methods may be attributed to two factors. The first, which we refer to as     ``grouping'' is the idea that some correct subsets of series are chosen to form new middle-level series. This is the argument commonly made to support clustering-based hierarchy construction \citep[see \textit{e.g.}][]{liForecastReconciliationApproach2019, pangHierarchicalElectricityTime2022, matteraImprovingOutofSampleForecasts2023}.  
%The idea that often motivates time series clustering is these ``groupings'' be composed of similar series. 
The second factor, which we refer to as the ``structure'' of the hierarchy, includes the number of middle-level series, the depth of the hierarchy, and the distribution of group sizes in the middle layer(s). Evidence showing that clustering within a forecast reconciliation framework leads to improved forecast accuracy, does not disentangle contributions from these two factors. Indeed, clustering methods may only work in so far as they generate a larger number of base forecasts. This argument would be consistent with the interpretation of reconciliation as a forecast combination of ``direct'' and ``indirect'' forecasts (\citealp{hollymanUnderstandingForecastReconciliation2021}), since more middle-level series implies a greater number of indirect forecasts in the combination. This leads to our second research question:

\begin{researchquestion}
    Should the improved accuracy of clustering-based methods be attributed to grouping together similar time series, or to the structure of the hierarchy? %including the number of middle level series?
\end{researchquestion}

To investigate this question, we devise the following approach. We take a hierarchy found using a certain clustering method (or even the natural hierarchy), and then randomly permute the bottom level series (\textit{i.e.}, the leaf nodes of the hierarchical tree). Multiple new ``twin'' hierarchies are formed with an identical structure to the original hierarchy, but with permuted leaves. In this way, we keep the \text{hierarchical structure} fixed, but alter how series are combined. This method can be thought of as an informal ``permutation type'' test \citep{welch1990}. Our main finding is that hierarchies constructed using clustering methods do not significantly outperform their random ``twins'', leading to the conclusion that the driver of forecast improvement is the enlarged number of series in the hierarchy and/or its structure, rather than similarities between the time series.

%Despite the intuitive appeal of clustering analysis, it is not straightforward to conclude that clustering itself is the primary contributor to the performance improvement illustrated in the literature. Except the factor of potentially accurate base forecasts of new series introduced by clustering, another factor  is the ``enriched structure'', meaning that a larger set of series is optimally combined to obtain the reconciled forecasts, leading to reduced model uncertainty and data uncertainty from forecast combination perspective (\citealp{hollymanUnderstandingForecastReconciliation2021}). The following question also motivates our research:
%\begin{researchquestion}
%Which factor, enriched structure or clustering, is the primary contributor to the performance improvement illustrated in clustering-based reconciliation approaches?
%\end{researchquestion}

%To address this question, we propose a random hierarchy construction approach. By randomly constructing ``twin'' hierarchies with the same tree structure but different permutations of bottom-level series as a contrast to a clustering-based hierarchy, we are able to isolate the influence of clustering algorithms. This allows us to inspect the individual contributions of the enriched structure and clustering technique to forecast reconciliation performance.

Finally, from a practical perspective, we investigate the role of forecast combination in cluster-based hierarchical forecasting. With multiple hierarchies available and inspired by the forecast combination literature (\citealp{wangForecastCombinations50year2022}), we consider the last research question

\begin{researchquestion}
    Does an equally-weighted combination of reconciled forecasts derived from multiple hierarchies improve forecast reconciliation performance?
\end{researchquestion}

Note that forecast combination here differs from that of \cite{hollymanUnderstandingForecastReconciliation2021}, in that our approach averages not only different coherent forecasts, but also across hierarchies with completely different middle level series. This is possible since only coherent bottom and top level forecasts are averaged and evaluated.

In summary, this paper presents four main contributions:

\begin{itemize}
  \item We introduce a novel hierarchical forecast reconciliation framework centered on hierarchy construction.  Within this framework, we introduce and compare three distinct approaches: cluster-based hierarchies, hierarchies based on random permutation, and  forecast combinations across different hierarchies.
  \item In contrast to existing literature that often focuses on a single clustering technique, our study systematically investigates the effectiveness of various time series clustering implementations. This investigation involves the incorporation of four time series representations, two distance measures, and two clustering algorithms.
  \item We conduct experiments using two empirical datasets - the Australian tourism dataset and the U.S. cause-of-death mortality dataset as well as a synthetic dataset. The results allow for a comparison of different approaches to constructing hierarchies.
  \item By constructing random hierarchies through permutation of leaf nodes, we show that the hierarchical structure is the primary contributor to improvements in forecast reconciliation performance, rather than the grouping of similar bottom level series.

\end{itemize}

The rest of the paper is organized as follows. Section~\ref{sec:methodology} describes the trace minimization reconciliation methods employed and the clustering-based hierarchical time series augmentation techniques considered. Section~\ref{sec:cluster} first introduces the two datasets used, and then investigates RQ1, in particular the performance of cluster hierarchy compared to natural hierarchy and two-level hierarchy. Section~\ref{sec:permutation} introduces the novel permutation approach, followed by the investigation of RQ2 via evaluating the performance of natural hierarchy and the best performing cluster hierarchy compared to their respective random twins. To avoid the concern that clusters found in the empirical datasets are spurious, a simulation study is considered in Section~\ref{sec:simulation}. Section~\ref{sec:combination} covers the forecast combination approach raised in RQ3. Section~\ref{sec:conclusion} concludes this paper with discussions on the findings and outlines future research directions. 
Data and code for reproducing the results in this paper are available at \url{https://github.com/AngelPone/project_hierarchy}.

%Section~\ref{sec:related_work} reviews related work in time series clustering, forecast reconciliation and clustering-based forecast reconciliation. In Section~\ref{sec:method} we introduce the proposed framework, elaborating on three approaches: cluster hierarchies, random hierarchies and combination hierarchies. The simulation study is presented in Section~\ref{sec:simulation}, while Section~\ref{sec:emp} details empirical experiments conducted on the Australian tourism dataset and U.S. mortality dataset. Finally, Section~\ref{sec:conclusion} concludes this paper with discussions on the findings and outlines future research directions.

\section{Methodology}\label{sec:methodology}

\subsection{Hierarchical forecasting and reconciliation methods}

%We propose a general hierarchical time series augmentation framework, in which new time series are constructed by aggregating subsets of bottom-level series. 
%Such subsets can be formed according to  results of time series clustering, attributes of the time series or other expert knowledge.

Hierarchical data can be characterized as being made up of bottom level series and their aggregates. In general, we consider a hierarchy with $n$ time series stacked into a vector $\boldsymbol{y}_t$. Let $\boldsymbol{b}_t$ denote the $m$ bottom level series, $\boldsymbol{a}_t$ denote the top level series, and $\boldsymbol{c}_t$ denote $k$ middle level series. The top level and bottom level will be common to all hierarchies we consider, and are augmented by middle level series. These middle level series can be formed according to attributes of the time series or in a data-driven fashion using time series clustering. The series are linked through an $(m+k+1)\times m$ summing matrix

%$\boldsymbol{S}$, which can be decomposed into an identity matrix $\boldsymbol{I}_m$ and two aggregation matrices $\boldsymbol{A}$ and $\boldsymbol{C}$ consisting of $0$ and $1$, such that, 
\[
  \boldsymbol{y}_t = \boldsymbol{S}\boldsymbol{b}_t = \begin{bmatrix}
    \boldsymbol{A} \\\boldsymbol{C} \\ \boldsymbol{I}_m 
  \end{bmatrix}  \boldsymbol{b}_t = \begin{bmatrix}
      \boldsymbol{a}_t \\ \boldsymbol{c}_t \\\boldsymbol{b}_t
  \end{bmatrix},
\]
where, $\boldsymbol{C}$ consists of zeros and ones that encode the aggregation, \textit{i.e.}, $c_{ij}=1$ if bottom level series $j$ is included in middle level series $i$, and zero otherwise. The top level series aggregates all bottom levels, \textit{i.e.}, $\boldsymbol{A} = \mathbf{1}_{1\times m}$, which is a row of ones.

%where $\boldsymbol{A}$ and  $\boldsymbol{C}$ represent the mappings from bottom-level time series to the given aggregated time series and constructed time series, respectively. 
%Each row of $\boldsymbol{C}$ corresponds to a subset of bottom-level series. The values $c_i, i=1,\dots,m$, indicate whether the $i$-th bottom-level series belongs to the subset.
% In the case of a two-level hierarchy, we simply have $\mathbf{A} = \mathbf{1}_{1\times m}$.
%We consider a two-level given hierarchy throughout this paper, where $\mathbf{A} = \mathbf{1}_{1\times m}$.

%\subsection{Trace minimization forecast reconciliation approach}

%Forecast reconciliation is a statistical technique aimed to improve the overall forecasts accuracy across different series in a hierarchical setting. It is a post-forecasting step that ensures the coherence of forecasts for all series in the hierarchy. Well-recognized regression types of reconciliation include the Ordinary Least Square (OLS) method proposed by \cite{hyndmanOptimalCombinationForecasts2011}, and the trace minimization (MinT) method proposed by \cite{wickramasuriyaOptimalForecastReconciliation2019}. 
%Forecast reconciliation is a hierarchical forecasting which first generates base forecasts for all time series in the hierarchy, and then reconciles the base forecasts via regression (\citealp{hyndmanOptimalCombinationForecasts2011}) or trace minimization (MinT, \citealp{wickramasuriyaOptimalForecastReconciliation2019}).
%It is well-known for the ability to produce coherent and more accurate forecasts, which has been exploited in several perspectives, such as forecast combination (\citealp{hollymanUnderstandingForecastReconciliation2021}) or projection (\citealp{panagiotelisForecastReconciliationGeometric2021}).
%In this subsection, we describe the process of generating coherent forecasts using the MinT approach, which will be employed throughout this paper.

% Consider a given hierarchy with $n$ time series with $m$ of them at the bottom level.  Let vectors $\boldsymbol{b}_t$, $\boldsymbol{a}_t$, and $\boldsymbol{y}_t$ represent observations  at time $t$ of the bottom level, the aggregated levels (\textit{i.e.}, middle and top levels), and the entire hierarchy, respectively.
% The vectors are linked through an $n\times m$ summing matrix $\boldsymbol{S}$, which can be decomposed into an identity matrix $\boldsymbol{I}_m$ and a constraint matrix $\boldsymbol{A}$ consisting of $0$ and $1$, such that, 
% \[
%   \boldsymbol{y}_t = \boldsymbol{S}\boldsymbol{b}_t = \begin{bmatrix}
%     \boldsymbol{A} \\ \boldsymbol{I}_m 
%   \end{bmatrix}  \boldsymbol{b}_t = \begin{bmatrix}
%       \boldsymbol{a}_t \\ \boldsymbol{b}_t
%   \end{bmatrix},
% \]
% where $\boldsymbol{A}$ represents the mapping from bottom-level time series to aggregated-level time series. In the case of a two-level hierarchy, we simply have $\mathbf{A} = \mathbf{1}_{1\times m}$.

%To produce $h$-step-ahead forecasts based on $T$ historic observations, we would first produce $h$-step-ahead unreconciled (or ``base'') forecasts $\hat{\boldsymbol{y}}_{T+h}$. Note that there are different ways to compute such forecasts. In this paper, we adopt the Exponential Smoothing (ETS) method, a well-known univariate forecasting model widely used in both research and practice \citep{ForecastingExponentialSmoothing}. 
%The \code{forecast} (\citealp{forecast}) package in {R} (\citealp{R}) is implemented to produce base forecasts for each time series in the hierarchy.
%Subsequently, the MinT reconciliation method is applied to produce coherent forecasts:
%\[
%    \tilde{\boldsymbol{y}}_{T+h} = \boldsymbol{S}(\boldsymbol{S}'\boldsymbol{W}_h^{-1}\boldsymbol{S})^{-1}\boldsymbol{S}'\boldsymbol{W}_h^{-1}\hat{\boldsymbol{y}}_{T+h},
%\]
%where $\boldsymbol{W}_h$ is the covariance matrix of $h$-step-ahead forecast errors. The shrinkage method is employed to obtain the estimator of $\boldsymbol{W}_h$ \citep{wickramasuriyaOptimalForecastReconciliation2019}. %is considered throughout this paper.
%Several estimators of $\boldsymbol{W}_h$ were proposed in \cite{wickramasuriyaOptimalForecastReconciliation2019}. %such as the OLS estimator, variance scaling, structural scaling estimator, and shrinkage estimator. 
% due to its ability to account for covariance in forecast errors.

For the purposes of this paper, all forecasting is carried out as a two-step process. First, so-called \textit{base} forecasts are produced for all series in the hierarchy and stacked into an $n$-vector $\hat{\bm{y}}$, where subscripts are suppressed for brevity. For base forecasts, we use the Exponential Smoothing (ETS) method \citep{ForecastingExponentialSmoothing}, implemented using the \code{forecast} (\citealp{forecast}) package in {R} (\citealp{R}). Alternative methods such as ARIMA models were also considered, but this choice had very little impact on the overall conclusions.

The base forecasts generated in this first step, do not have the property that bottom level forecasts add up to forecasts of the aggregates, \textit{i.e.}, they are \textit{incoherent}. Therefore, forecast reconciliation is used as a post-forecasting step to ensure coherence of forecasts for all series in the hierarchy. In general, reconciliation takes the form of projecting the base forecasts as
\[
    \tilde{\boldsymbol{y}} = \boldsymbol{S}(\boldsymbol{S}'\boldsymbol{W}_h^{-1}\boldsymbol{S})^{-1}\boldsymbol{S}'\boldsymbol{W}_h^{-1}\hat{\boldsymbol{y}},
\]
where $\tilde{\boldsymbol{y}}$ are the \textit{reconciled} forecasts and $\boldsymbol{W}_h$ is the covariance matrix of $h$-step-ahead forecast errors. For reconciliation, we use the MinT method \citep{wickramasuriyaOptimalForecastReconciliation2019}, which involves plugging in a shrinkage estimator for $\boldsymbol{W}_h$. Although alternatives were considered, including approaches that assume $\boldsymbol{W}_h$ is diagonal or an identity matrix \citep{hyndmanOptimalCombinationForecasts2011}, this choice has little impact on our main conclusions.

%\subsection{Grouping-based forecast reconciliation}
%\subsection{Augmenting hierarchical time series through clustering} 
\subsection{Time series clustering} 
\label{sec:clustering}

%\textbf{HOW TO AUGMENT THE HIERARCHY}
%\subsection{Cluster hierarchy construction approach}

To the best of our knowledge, four studies have attempted to improve forecast accuracy in a reconciliation setting by constructing middle levels of the hierarchy using time series clustering.
\cite{pangHierarchicalElectricityTime2018} detect consumption patterns of electricity smart meter data based on X-means clustering algorithm, while \cite{pangHierarchicalElectricityTime2022} propose several alternative clustering methods to group similar electricity and solar power time series. 
\cite{liForecastReconciliationApproach2019} apply agglomerative hierarchical clustering to cause-of-death time series, and
\cite{matteraImprovingOutofSampleForecasts2023} utilize Partition Around Medoids algorithms to unveil underlying structures in stock price indexes. 
However, these studies are limited in scope as they focus on a few  clustering techniques. 
%\todo{Should we move to intro?}
%construct hierarchies through clustering algorithms and 
Inspired by the comprehensive overview of time series clustering by \cite{aghabozorgiTimeseriesClusteringDecade2015a}, we consider various approaches based on three key components, namely \textit{time series representations}, \textit{distance measures}, and \textit{clustering algorithms}. %Our framework for creating a clustering approach is as follows: we first choose a time series representation, which can be the raw time series, the in-sample forecast error, features of the time series, or features of the in-sample forecast error. We then select a distance measure, either Euclidean distance or dynamic time warping. Once the representation and the distance measure are determined, we choose a clustering algorithm, either $k$-medoids or agglomerative hierarchical clustering, to form clusters in the dataset. In total, we consider 12 different clustering approaches in our experiments.
%, and evaluation measures.

% Time series clustering is an essential technique that finds applications across various domains, including finance (\citealp{DIAS2015852}), energy (\citealp{pangHierarchicalElectricityTime2018}), robotics (\citealp{gopalapillaiExperimentationAnalysisTime2014}), and others. It is commonly used to identify similar patterns within time series datasets, aiding in applications such as anomaly detection, pattern discovery, and recommendation.

\paragraph{\textbf{Time series representations}}

The time series representation refers to the object that acts as an input for time series clustering. %that provide diverse perspectives on the same dataset.
% Considering the impracticality of exploring every conceivable time series representation, we focus on four key ones: raw time series without any transformation (hereafter referred to as ``time series''), in-sample one-step-ahead forecast error, features of time series, and features of in-sample one-step-ahead forecast error.
Our first candidate for the representation is raw time series itself, due to its simplicity and broad applicability.
%However, a crucial element contributing to the success of the MinT method lies in estimating the covariance matrix of base forecast errors based on in-sample one-step-ahead forecast error. 
We also consider the in-sample one-step-ahead forecast error as a representation of the time series, since a key step in MinT reconciliation is to estimate the $\boldsymbol{W}_h$ matrix. %to examine how the structure within the error series influences reconciliation performance. 
It is important to note that raw time series and in-sample error representations are standardized to eliminate the impact of scale variations. %This is done by normalizing each series by subtracting its mean and then dividing by the standard error.

A potential shortcoming to using the time series or forecast error as a representation is their high dimensionality, which is equal to the sample size of the training data. To address this concern, low dimension summaries of ``features'' can be considered. Features have been used in the context
%Features, widely employed in capturing time series characteristics, play a pivotal role in various time series applications, including 
time series clustering by \cite{tianoFeatTSFeaturebasedTime2021}, and for forecasting by \cite{wangUncertaintyEstimationFeaturebased2022} and \cite{ liFeaturebasedIntermittentDemand2023}. 
We consider features of both the raw time series and the in-sample forecast error as representations. After filtering out the features that are constant across all series, we select $56$ features.\footnote{The list and descriptions of features are available in the online GitHub repository.} These time series features are calculated by the \code{tsfeatures} (\citealp{tsfeatures}) package in R. %have been applied in several feature-based forecasting studies, \textit{e.g.}, \cite{montero-mansoFFORMAFeaturebasedForecast2020}.
To the best of our knowledge, we are the first to utilize in-sample forecast error and time series features as representations in the context of forecast reconciliation.
%\sout{These representations allow us to gain insights into the diverse aspects of hierarchical time series data}.

\paragraph{\textbf{Distance measures}}

All clustering algorithms we consider require a distance to be defined between the objects that act as inputs to the algorithm.
%Distance measures serve as crucial tools for assessing the similarity between two series, which in turn significantly influence the clustering results.
We consider two widely applied distance measures: Euclidean distance and dynamic time warping (DTW). When employing Euclidean distance, dimension reduction is necessary due to the curse of dimensionality. %, which acts as the dimension over which Euclidean distance is computed in our applications. 
%Failure to undertake dimension reduction may result in undesirable clustering outcomes due to the curse of dimensionality. 
To address this, we perform Principal Component Analysis (PCA), extracting the first few principal components that collectively explain at least 80\% of the variance within the representations.

Unlike Euclidean distance, DTW is not as sensitive to the curse of dimensionality (\citealp{sakoeDynamicProgrammingAlgorithm1978}). Instead of performing one-to-one point comparisons, DTW accommodates time series of varying lengths through many-to-one comparisons. This flexible approach allows for the recognition of time series with similar shapes, even in the presence of signal transformations such as shifting and/or scaling.

\paragraph{\textbf{Clustering algorithms}}
%\sout{Clustering algorithms, such as partitioning, hierarchical, grid-based, and model-based approaches,  identify clusters by optimizing specific objective functions} \citep{aghabozorgiTimeseriesClusteringDecade2015a}. %, or applying heuristics. %These functions are calculated based on chosen time series representations and distance measures.
In this paper, we focus on two clustering algorithms, namely $k$-medoids and agglomerative hierarchical clustering. These algorithms are implemented using the \texttt{cluster} (\citealp{cluster}) package in R. {The $k$-medoids algorithm aims to minimize the total distance between all observations within a cluster and their respective cluster median.} This is implemented using the partitioning around medoids (PAM) method (\citealp{PartitioningMedoidsProgram1990}).
%Unlike $k$-means which employs the mean vector of samples as the cluster center, $k$-medoids selects one observation within each cluster as the center. 
Following the recommendation of \cite{PartitioningMedoidsProgram1990}, we determine the optimal number of clusters using the average silhouette width (ASW), which is a commonly used index in cluster validation \citep[see \textit{e.g.},][]{shutaywi2021silhouette}.
%ASW assesses the quality of clustering results by measuring the proximity of samples within a cluster compared to neighboring clusters.
%Given a clustering result, the silhouette width for the $i$th sample is calculated as 
%\[
%  SW(i) = \frac{b(i)-a(i)}{\max\{a(i), b(i)\}},  
%\]
%where $a(i)$ represents the average distance of the $i$th observation to others in the same cluster, and $b(i)$ is the average distance to observations in the cluster nearest to observation excluding the cluster that observation $i$ is assigned to. A higher silhouette width indicates greater proximity to observations within the same cluster than to those in neighboring clusters. The ASW is the average of silhouette width across all observation. We select the clustering result that maximizes the ASW by iterating over all possible numbers of clusters. However, ASW has the limitation of being undefined when there is only one cluster. %\textbf{Gap statistics (\citealp{tibshiraniEstimatingNumberClusters2002}) can be used to address this limitation, but it is incompatible with DTW.}

Agglomerative hierarchical clustering initializes each observation in its own cluster, and then merges the nearest two clusters in a stepwise fashion until all observations form a single cluster. We employ Ward's linkage (\citealp{murtaghWardHierarchicalAgglomerative2014a}) which defines the nearest clusters as those that minimize the increase in within-cluster variance at each step. Applying hierarchical clustering to $m$ bottom-level series results in a binary hierarchical tree with ($2m-1$) nodes, all of which are retained as middle level series.

\begin{figure}[h!]
    \centering
    \vspace{0.2in}
    \includegraphics[width=0.46\textwidth]{figures/pamcluster.pdf}
    \hspace{1cm}
    \includegraphics[width=0.46\textwidth]{figures/aggcluster.pdf}
    \caption{\label{fig:cluster_example}Example clustering results of two clustering algorithms. Left panel displays example for $k$-medoids algorithm, and right panel displays example for agglomerative hierarchical clustering algorithm.}
\end{figure}
Figure~\ref{fig:cluster_example} illustrates two examples of hierarchies generated by $k$-medoids and agglomerative clustering algorithms, on the left and right panels respectively. The corresponding two-level hierarchy would have four bottom-level series and one top-level series. % \sout{These examples highlight the distinct behaviors of the two algorithms.}
Note that, $k$-medoids constructs a simple hierarchy with a single middle level, while hierarchical clustering generates multiple nested middle levels. 
In general, $k$-medoids produces a hierarchy with fewer middle levels and middle level series compared to hierarchical clustering. As the number of bottom-level series increases, these differences become increasingly pronounced, with potential implications for forecast reconciliation.

In summary, we employ 12 time series clustering approaches which are derived from combinations of four time series representations, two distance measures, and two clustering algorithms. The names and details of these approaches are listed in Table~\ref{tab:P3_methods}. Note that all methods using DTW, have either the raw time series or forecast errors as representations, since DTW is not compatible with time series features. \\

\begin{table}[h!]
\caption{\label{tab:P3_methods} Details of the 12 clustering approaches considered.}
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{rcrrr}
    \toprule
    Approach & Dimension reduction & Representation & Distance measure & Clustering algorithm  \\ \midrule
    TS-EUC-ME &  Yes & Time series  & Euclidean & $k$-medoids \\
    ER-EUC-ME & Yes& In-sample  error  & Euclidean & $k$-medoids \\
    TSF-EUC-ME & Yes& Time series features  & Euclidean & $k$-medoids \\
    ERF-EUC-ME & Yes& In-sample error features  & Euclidean & $k$-medoids \\
    TS-EUC-HC & Yes& Time series  & Euclidean & Hierarchical   \\ 
    ER-EUC-HC & Yes& In-sample error  & Euclidean & Hierarchical   \\ 
    TSF-EUC-HC & Yes& Time series features  & Euclidean & Hierarchical   \\ 
    ERF-EUC-HC & Yes& In-sample error features  & Euclidean & Hierarchical  \\
    TS-DTW-ME & No& Time series  & DTW & $k$-medoids \\
    TS-DTW-HC & No& In-sample error  & DTW & Hierarchical  \\
    ER-DTW-ME & No& Time series  & DTW & $k$-medoids \\
    ER-DTW-HC & No& In-sample error  & DTW & Hierarchical 
     \\\bottomrule
\end{tabular}}
\end{table}

%The four time series representations need to perform dimension reduction when using Euclidean distance. 
%Combing the four dimension reduced representations with two clustering algorithms results in the first eight clustering approaches. 
%The first eight approaches require dimension reduction via PCA, while the last four approaches use either the raw time-series or forecast errors, 
%must take temporal data as an input.
%the last four approaches 
%Note that features of raw time series and in-sample forecast error are not temporal data, thus incompatible with DTW. Therefore, the last four approaches are derived from combinations of two temporal representations (\textit{i.e.}, raw time series and in-sample error), DTW, and two clustering algorithms.

\section{Improving forecast performance via hierarchy augmentation}\label{sec:cluster}
\subsection{Data description}

We conduct our experiments on two empirical datasets throughout this paper. The first is the monthly Australian domestic tourism dataset, covering the period from January 1998 to December 2016.\footnote{Please refer to Section 4 of \cite{wickramasuriyaOptimalForecastReconciliation2019} for an in-depth explanation of this dataset.} The data is recorded as ``visitor nights'', representing the total number of nights spent by Australians away from home. In this dataset, the total visitor nights of Australia is geographically disaggregated into seven states and territories, which are further divided into $27$ zones, and then into $76$ regions. Additionally, each regional-level series is divided by four travel purposes. Overall, this dataset comprises a total of $555$ time series with $304$ of those at the bottom level. In the case of tourism data, the first two or three letters of the series name indicate geographical zones or regions, and the last three denote travel purposes. For example, ``\textit{AAAHol}'' represents the visitor nights spent for holiday in the ``Sydney'' region.

>>>>

% \usepackage{natbib}
\usepackage{amsfonts,amsmath,bbm,bm,xcolor,booktabs,hyperref,amsthm}
\usepackage{geometry}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{subfig}
\geometry{a4paper,scale=0.8}
\usepackage{setspace}
\setstretch{1.5}
\let\code=\texttt
\let\proglang=\textsf
\setcounter{MaxMatrixCols}{20}
\usepackage{enumitem}
% \usepackage[pagewise]{lineno}

@@

@@

@@

\section*{Acknowledgements}

Bohan Zhang is supported by the international joint doctoral education fund of Beihang University. Han Li is supported by the Australian Research Council Discovery Projects funding scheme (DP220100090).

\section*{Declaration of competing interest}

The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
